BarnebyLives: an R package to create herbarium specimen labels and clean spreadsheets

### 2025-03-26 manuscript/BarnebyLives An R package to create and qc herbarium labels.rmd
ordered coll nos

```{r install package}
## devtools::install_github('sagesteppe/BarnebyLives')
```

@@

```{r, eval = T}
## this SHOULD be a sh chunk, so launch in a shell! I just want to be sure you can see the code. 
## and eval = TRUE will be respected. 
user: pwd
/media/steppe/hdd/Barneby_Lives-dev/manuscript/labels/raw
# this will have both your script and a dir full of raw labels with a users name
# now launch the bash script with: 
bash render_labels.sh collector='Reed'
# or chmod it etc!
# et voila! data will be saved up on dir in a 'final' directory. 
```

@@

time_label_gen <- readRDS('./data/processed/time_label_gen')[['elapsed']]
time_4perpage <- gsub('user.*$', '', readLines('./labels/4labsperpage.txt'))[1]
time_final <- gsub('user.*$', '', readLines('./labels/process2final.txt'))[1]

@@

Most of the local run time is attributable to the spatial (`r sum(tt_local[tt_local$Module=='Geospatial', 'elapsed'])`sec), and taxonomic operations (`r sum(tt_local[tt_local$Module=='Taxonomic', 'elapsed'])`sec), while formatting data for labels took `r sum(tt_local[tt_local$Module=='Style', 'elapsed'])`sec. 
The spell check of the scientific name accounted for nearly all of the time (`r sum(tt_local[tt_local$Function=='spell_check', 'elapsed'])`sec) spent performing local taxonomic operations.
The generation of labels consumed around `r minutemen(time_label_gen)` minutes (`r time_label_gen`sec) for the rendering, `r time_4perpage`sec to combine individual labels four per single sheet of landscape orientated paper, and `r time_final`sec to combine the `r length(list.files('labels/processed'))` sheets to a single Portable Document Format (PDF). 
The total computer run time for processing these `r nrow(data)` specimens was `r minutemen(sum(time_trials$elapsed))` minutes.

@@

## Family Level Results
# these species were found by POW, and according to POW, the species are in the f
# following families. 
family_incongruence <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(POW_Family != 'NOT FOUND') %>% 
  filter(POW_Family != Family) %>% 
  select(Collection_number, Full_name, Family, POW_Family) %>% 
  mutate(Situation = 
    case_when(
      Family %in% c('Hydrophyllaceae', 'Namaceae') ~ 'Preferred', 
      Family %in% c('Athyriaceae') ~ 'Outdated',
      Family %in% c('Saxifragaceae', 'Plantaginaceae') ~ 'Submitter_Incorrect', 
      Family %in% c('Bataceae', 'Ericaceae') ~ 'Internal_Error', 
      .default = 'Typo'
    ))

# how many of my genera were mis-spelled??
genera_mispelled <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(POW_Genus == 'NOT FOUND') %>% 
  select(Collection_number, Full_name, Genus, POW_Genus)

>>>>

```{r install package, eval = F}
## devtools::install_github('sagesteppe/BarnebyLives')
```

@@

```{r combine labels, eval = F}
## this SHOULD be a sh chunk, so launch in a shell! I just want to be sure you can see the code. 
## and eval = TRUE will be respected. 
user: pwd
/media/steppe/hdd/Barneby_Lives-dev/manuscript/labels/raw
# this will have both your script and a dir full of raw labels with a users name
# now launch the bash script with: 
bash render_labels.sh collector='Reed'
# or chmod it etc!
# et voila! data will be saved up on dir in a 'final' directory. 
# we timed it like this:
\time -o ../process2final.txt  bash render_labels.sh collector='Reed'
```

@@

time_label_gen <- readRDS('./data/processed/time_label_gen')[['elapsed']]
time_final <- gsub('user.*$', '', readLines('./labels/process2final.txt'))[1]

@@

Most of the local run time is attributable to the spatial (`r sum(tt_local[tt_local$Module=='Geospatial', 'elapsed'])`sec), and taxonomic operations (`r sum(tt_local[tt_local$Module=='Taxonomic', 'elapsed'])`sec), while formatting data for labels took `r sum(tt_local[tt_local$Module=='Style', 'elapsed'])`sec. 
The spell check of the scientific name accounted for nearly all of the time (`r sum(tt_local[tt_local$Function=='spell_check', 'elapsed'])`sec) spent performing local taxonomic operations.
The generation of labels consumed around `r minutemen(time_label_gen)` minutes (`r time_label_gen`sec) for the rendering, and an additional `r time_final`sec to combine the `r length(list.files('labels/processed'))` sheets to a single Portable Document Format (PDF). 
The total computer run time for processing these `r nrow(data)` specimens was `r minutemen(sum(time_trials$elapsed))` minutes.

@@

## Family Level Results
# these species were found by POW, and according to POW, the species are in the f
# following families. 
family_incongruence <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(POW_Family != 'NOT FOUND') %>% 
  filter(POW_Family != Family) %>% 
  select(Collection_number, Scientific_name, Family, POW_Family) %>% 
  mutate(Situation = 
    case_when(
      Family %in% c('Hydrophyllaceae', 'Namaceae') ~ 'Preferred', 
      Family %in% c('Athyriaceae') ~ 'Outdated',
      Family %in% c('Saxifragaceae', 'Plantaginaceae') ~ 'Submitter_Incorrect', 
      Family %in% c('Bataceae', 'Ericaceae') ~ 'Internal_Error', 
      .default = 'Typo'
    ))

# how many of my genera were mis-spelled??
genera_mispelled <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(POW_Genus == 'NOT FOUND') %>% 
  select(Collection_number, Scientific_name, Genus, POW_Genus)
